---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
    theme: readable
    highlight: tango
    toc: yes
    toc_depth: 2
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```
## Packages 
```{r}
library(ggplot2)
library(lubridate)
library(dplyr)
library(lattice)
library(xtable)
library(impute)
```


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
data$date <- strptime(data$date, format = "%Y-%m-%d")
data$day<- as.integer(difftime(data$date, data$date[1], units = "days"))
```


## What is mean total number of steps taken per day?

I Summary in data.frame *df*,  total, mean and median steps by day.

```{r}
summary_steps <- data %>% group_by(day) %>% 
    summarise(Total = sum(steps, na.rm = T), Media=mean(steps, na.rm = T), 
              Mediana = median(steps, na.rm = T))
```


```{r}
ggplot(summary_steps, aes(day, Total)) + ylab("Steps") + xlab("Day")+
    geom_bar(stat = "Identity")
```

```{r, results='asis'}
xt <- xtable(summary_steps, caption = "Table 1. Mean and median number of steps taken each day", 
             align = "ccccc", digits = 2)
print(xt, type="html", html.table.attributes = 'align="center", width=40%, 
                                                frame="below", rules="rows"')
```

## What is the average daily activity pattern?

```{r}
df <- data %>% group_by(interval) %>% summarise(media=mean(steps, na.rm = T))
(max_steps <- df[which.max(df$media),])
```

```{r}
ggplot(df, aes(interval, media)) + geom_line(size=0.7, color = "blue") + xlab("Minutes") + 
    geom_vline(xintercept = max_steps$interval[1], linetype="dashed", color = "red") + 
    ylab("Average steps daily") + geom_text(label="Max Interval = 835", x = 1200, y = 200) 
```


## Imputing missing values

```{r, results='asis'}
na <- data %>% mutate_all(is.na) %>% summarise_all(sum)
xt <- xtable(na, caption = "Table 2. Total missing values", 
             align = "ccccc", digits = 2)
print(xt, type="html", html.table.attributes = 'align="center", width=40%, 
                                                frame="below", rules="rows"')
```

```{r, results='hide'}
df1 <- data %>% mutate(date = as.numeric(date))
df_impute <- as.data.frame(impute.knn(as.matrix(df1))$data)
```

```{r}
summary_impute <- df_impute %>% group_by(day) %>% 
    summarise(Total = sum(steps, na.rm = T), Media=mean(steps, na.rm = T), 
              Mediana = median(steps, na.rm = T))

```

```{r}
ggplot(summary_impute, aes(day, Total)) + ylab("Steps") + xlab("Day") +
    geom_bar(stat = "Identity", position=position_dodge(width=.9))
     
```

```{r, results='asis'}
xt <- xtable(summary_impute,,align = "ccccc", digits = 2, 
             caption = "Table 3. Mean and median number of steps taken each day with missing values imputing")
print(xt, type="html", html.table.attributes = 'align="center", width=40%, 
                                                frame="below", rules="rows"')
```

### Differences
The mean of total increase imputing the missing values. some days with NAs change the values, however the mean of means is equal.  
```{r}
na_compare <- bind_rows("with_NA" = summary_steps, "impute_NA" = summary_impute, 
                        .id = "set") %>% mutate(set = as.factor(set))
    
ggplot(na_compare, aes(day, Total, fill=set)) + ylab("Steps") + xlab("Day") +
    geom_bar(stat = "Identity", position=position_dodge())
```

```{r}
with(na_compare, tapply(Media, set, mean, na.rm=T))
with(na_compare, tapply(Total, set, mean, na.rm=T))
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
data$wd <- as.factor(weekdays(data$date))
levels(data$wd) <- c("Weekend", rep("Weekday", 4), "Weekend", "Weekday")
df <- data %>% group_by(interval, wd) %>% summarise(media=mean(steps, na.rm = T))
xyplot(media~interval|wd, df, type="l", layout=c(1,2), ylab = "Steps", xlab = "Minutes")
```

